---
Title: "color-model-hlcam"
author: "David Morrison"
date: "07/08/2017"
---
# High Luminance Color Appearance Model 
This document aims to develop an improved color appearance model that can be used for the accurate analysis of images for artists and painters.
It is based on the 2009 research into extended luminance color appearance models beyond CIECAM02.

## References
* Kim, M., Weyrich, T. , Kautz, L, 2009. **Modeling human color perception under extended luminance levels**, ACM 
Trans. Graph. 28 (3), http://reality.cs.ucl.ac.uk/projects/xlrcam/kim09xlrcam.pdf


```{r Initialise, echo=FALSE}
require(colorscience)
require(ggplot2)
require(dplyr)
source("Projects/color-model-hlcam/hlcamFunctions.R")

```
## Load the experimental data
Extract the data from the pdf at http://jankautz.com/publications/experimental_data.pdf
1. Copy the PDF's text to data_raw.txt
2. Run the node program 'node data_extract.js'

```{r loaddata}
all=read.csv(file="Projects/color-model-hlcam/data_perceptual.csv",header=TRUE)
```

## Explore the data
Luminance Range and Background Level has strongest influence on color perception.

Lighting Conditions
1. CCT - Correlated color termperature of illuminant in 'Kelvin. Candle (1850K), Incandescent Lamp (2400K), Warm CFL (3000K), Moonlight (4100K), Horizon daylight (D50 5000K), Vertical daylight (D55 5500K - D60 6000K), Midday (5780K), Overcast/LCD (D65 6500K), Clear blue poleward sky (15000K - 27000K)
2. Xw, Yw, Zw - XYZ for Adopted White in test illuminant, Yw is also the Peak Luminance given its white
3. Pb - Background lightness percentage = Yb / Yw * 100. Ranges from 0 to 100.
4. Xb, Yb, Zb - XYZ for Background

Physical measures
1. Xs, Ys, Zs - XYZ for sample under test illuminant

Perceptual measures
1. Hs - Hue quadrature. Varies from 0 to 400: redness (0) - yellowness (100) - greenness (200) - blueness (300) - redness (400)
2. Js - Lightness. A scale of 0–100 relative to the brightness of reference white
3. Ms - Colorfulness. An absolute scale of 0 to unlimited, but typically 0 to 100


```{r explore}


## confirm relationship between CCT and Xw, Yw, Zw
## CCT1 calc using McCamy's Approximation https://en.wikipedia.org/wiki/Color_temperature#Approximation
## inverse using Plankian locus Approximation https://en.wikipedia.org/wiki/Planckian_locus#Approximation
all$CCT1 = XYZtoCCT(all$Xw, all$Yw, all$Zw)
ggplot(data=all)+aes(x=CCT, y=CCT1)+geom_point()

## confirm relationship between Xw, Yw, Zw and Xb, Yb, Zb is just the Pb ratio 
all$Yb1 = all$Yw * all$Pb / 100
ggplot(data=all)+aes(x=Yb, y=Yb1, color=Ms)+geom_point()
all$Xb1 = all$Xw * all$Pb / 100
ggplot(data=all)+aes(x=Xb, y=Xb1, color=Ms)+geom_point()
all$Zb1 = all$Zw * all$Pb / 100
ggplot(data=all)+aes(x=Zb, y=Zb1, color=Ms)+geom_point()

## confirm any relationship between La and Yb, should be similar since the background takes up the most of the adaption field
ggplot(data=all)+aes(x=Yb, y=La, color=Pb)+geom_point()

## explore measures Xs, Ys, Zs
ggplot(data=all)+geom_jitter(aes(y=Xs,x="Xs"))+geom_jitter(aes(y=Ys,x="Ys"))+geom_jitter(aes(y=Zs,x="Zs"))+geom_jitter(aes(y=La,x="La"))+ theme(axis.title.x = element_blank())+ theme(axis.title.y = element_blank())

## explore perceptual measures Js, Ms, Hs
ggplot(data=all)+geom_jitter(aes(y=Js,x="Js"))+geom_jitter(aes(y=Ms,x="Ms"))+geom_jitter(aes(y=Hs,x="Hs"))+ theme(axis.title.x = element_blank())+ theme(axis.title.y = element_blank())
ggplot(data=all)+aes(x=Hs, y=Js, color=Ms)+geom_point()

all <- subset(all, select = -c(CCT1, Xb1, Yb1, Zb1))

```

## Perform Chromatic Adaption
Use the CEICAM02 Chromatic Adaption Algorithm called CAT02
Refer to https://en.wikipedia.org/wiki/CIECAM02

```{r ChromaticAdaption}

## convert (color sample) and (adopted white) to LMS space
LMSs <- XYZtoLMS(all$Xs,all$Ys,all$Zs)
LMSw <- XYZtoLMS(all$Xw,all$Yw,all$Zw)

## use CAT02 to adapt the sample from the adopted white illumant to a reference white used in CIECAM02
all$D <- calculateD(all$Ambient, all$La)
LMSsc <- LMStoLMSc(all$D, LMSs, LMSw)
LMSwc <- LMStoLMSc(all$D, LMSw, LMSw)

## perform post adaption conversion to cone responses using Hunt-Pointer-Esteves
LMSsp <- LMSctoLMSp(LMSsc)
LMSwp <- LMSctoLMSp(LMSwc)

## convert to absolute cone responses
LMSsr <- LMSptoLMSr(LMSsp, all$La)
LMSwr <- LMSptoLMSr(LMSwp, all$La)

## Calculate achromatic signals by averaging across cone responses in 40:20:1 ratio
As <- calculateA(LMSsr)
Aw <- calculateA(LMSwr)

## calculate the perceived lightness (0 to 100)
all$Js1 <- calcLightness(As, Aw, 1.7526)
ggplot(data=all)+aes(x=Js, y=Js1, color=Hs)+geom_point()

## Calculate the Hue Quadrature based on chromatic signals
ab <- calculateab(LMSsr)
all$Hs1 <- calcHueQ(ab[,1], ab[,2])
ggplot(data=all)+aes(x=Hs, y=Hs1, color=Hs)+geom_point()

all$Cs1 <- calcChroma(ab[,1], ab[,2])
ggplot(data=all)+aes(x=Ms, y=Cs1, color=Hs)+geom_point()

```
## Cone Response
Use the High Luminance Color Appearance Model
Refer to http://reality.cs.ucl.ac.uk/projects/xlrcam/kim09xlrcam.pdf

```{r ConeResponse}

## convert to absolute cone responses
nc = 0.57
Lp <- LMSp[,1]^nc / (LMSp[,1]^nc + all$La^nc)
Mp <- LMSp[,2]^nc / (LMSp[,2]^nc + all$La^nc)
Sp <- LMSp[,3]^nc / (LMSp[,3]^nc + all$La^nc)

## Calculate achromatic attributes by averaging across cone responses in 40:20:1 ratio
A <- (40 * Lp + 20 * Mp + Sp) / 61





## perform the chromatic adaption: convert from C -> D65 using Bradford method
mun_XYZ_D65 <- mun_XYZ_C %*% M_adapt_C_to_D65
all$D65X <- mun_XYZ_D65[,1]
all$D65Y <- mun_XYZ_D65[,2]
all$D65Z <- mun_XYZ_D65[,3]
Total <- all$D65X + all$D65Y + all$D65Z
all$CIEy=ifelse(Total==0, 0.32902, all$D65Y / Total)
all$CIEx=ifelse(Total==0, 0.31271, all$D65X / Total)
all$CIEY=all$D65Y

  
remove(M_adapt_C_to_D65,mun_XYZ_C,Total)
```

## Review and cleanup the data
Lets review the data with a jitter plot to see if we need to cleanup the data
Also lets add an ordered Hue scale and a numeric Hue scale that will help with later modelling. 
 

```{r Cleanup}

## Use a jitter plot to look for any outliers
ggplot(data=all)+geom_jitter(aes(y=D65X,x="D65X"))+geom_jitter(aes(y=D65Y,x="D65Y"))+geom_jitter(aes(y=D65Z,x="D65Z"))+scale_y_continuous(breaks=seq(-8,8))+ theme(axis.title.x = element_blank())+ theme(axis.title.y = element_blank())

## also later noticed a bad data point at H==7.5PB, V==0.2, C==20
## allx=all%>%filter(C>0,V==0.2|V==5|V==8)
## ggplot(data=allx,aes(y=D65X,x=H,colour=C))+geom_line(aes(group=C))+geom_point()+facet_wrap(~V,scales="free")

## Remove the obvious outliers
all <- all%>%filter(D65X>-0.2,D65Z>-0.2,D65Z<3,!((H=="7.5PB")&(V==0.2)&(C==20)))
mun_XYZ_D65 <- matrix(c(all$D65X,all$D65Y,all$D65Z), ncol=3)

## Check the cleanup
ggplot(data=all)+geom_jitter(aes(y=D65X,x="D65X"))+geom_jitter(aes(y=D65Y,x="D65Y"))+geom_jitter(aes(y=D65Z,x="D65Z"))+scale_y_continuous(breaks=seq(-8,8))+ theme(axis.title.x = element_blank())+ theme(axis.title.y = element_blank())

```

## Modelling the dataset
Our aim is to be able to predict the values of R, G, and B based on given values of HA, V and C. To simplify things we will actually develop models for D65X and D65Z then convert them to RGB with the known formulas. Since D65Y is very closely related to V we will only need to develop two non-linear models.
* D65X = fn1(HA,V,C)
* D65Z = fn2(HA,V,C)
* D65Y = fn3(V)

Lets start by looking at the dataset for D65X and D65Z, the two variables we are trying to predict. We will draw scatterplots, one scatterplot for each increasing value of V. The scatterplots will draw points with y=D65X against x=HA. We'll draw a line through all the datapoints with the same C. Key points to note are
*The overall relationship appears periodic with respect to HA from 0 to 360 (cf the color circle)
*As C nears 0 there is less influence from the HA until its a flat line (cf less Chroma means less Hue effect)
*As C increases, D65X reduces in the Green area (HA=180) and increases in the RedPurple area (HA=360)
*As C increases, D65Z reduces in the Yellow area (HA=90) and increases in the PurpleBlue area (HA=270)
*As V increases, both D65X and D65Z increase and the whole wave shifts up (cf more brightness)
*At low V there is small variation from the grey point, at high V high variation


```{r FirstView, fig.width=15, fig.height=15}
## create some statistics across all hues
stats <- all%>%filter(V> -1)%>%group_by(V,C)%>%summarise(n=n(),
          Xmin=min(D65X),Xmax=max(D65X),Xmean=mean(D65X),Xmed=median(D65X),Xmid=(Xmax+Xmin)/2,Xmag=(Xmax-Xmin)/2,
          Ymin=min(D65Y),Ymax=max(D65Y),Ymean=mean(D65Y),Ymed=median(D65Y),Ymid=(Ymax+Ymin)/2,Ymag=(Ymax-Ymin)/2,
          Zmin=min(D65Z),Zmax=max(D65Z),Zmean=mean(D65Z),Zmed=median(D65Z),Zmid=(Zmax+Zmin)/2,Zmag=(Zmax-Zmin)/2)


## view the data - all the data, one chart for each Value
ggplot(data=all%>%filter(C>-1,C<18),aes(y=D65Z,x=HA,colour=C))+geom_path(aes(group=C))+geom_point()+scale_x_continuous(breaks=seq(0, 360, by=360/8))+facet_wrap(~V)

## view the data - all the data one chart for each Value
ggplot(data=all%>%filter(C>-1),aes(y=D65Z,x=HA,colour=C))+geom_line(aes(group=C))+geom_point()+scale_x_continuous(breaks=seq(0, 360, by=360/8))+facet_wrap(~V)

## deeper examination where C==0
ggplot(data=all%>%filter(C==0),aes(y=D65X,x=V))+geom_line()+geom_point(color="blue")+geom_line(aes(y=D65Z,color="red"))+geom_point(aes(y=D65Z,color="red"))

## Model the mean greypoint data against V where C==0, ie X is a fn(poly(V,5)) and Z is a simple ratio of X
all$XZratio=all$D65Z/all$D65X
fit0X = lm(D65X~0+poly(V,5,raw=T),data=all%>%filter(C==0)); summary(fit0X)
fit0Z = lm(D65Z~0+poly(V,5,raw=T),data=all%>%filter(C==0)); summary(fit0Z)
fit0Y = lm(D65Y~0+poly(V,5,raw=T),data=all%>%filter(C==0)); summary(fit0Y)

## Store away the means of the grey points5
all$D65Xmean = predict(fit0X,all)
all$D65Zmean = predict(fit0Z,all)
all$D65Ymean = predict(fit0Y,all)
all$D65Xoffset = all$D65X - all$D65Xmean
all$D65Zoffset = all$D65Z - all$D65Zmean
all$D65Yoffset = all$D65Y - all$D65Ymean

## Model the standard deviation of the data against V where C==4. ie sd(X) is a fn(poly(V,3))
s <- all%>%filter(V>0,C==4)%>%group_by(V,C)%>%summarise(n=n(),Xsd=sd(D65Xoffset),Zsd=sd(D65Zoffset),Xmean=mean(D65Xoffset),Zmean=mean(D65Zoffset))
ggplot(data=s,aes(x=V,y=Xsd))+geom_line()          
fit0Xsd=lm(Xsd~poly(V,3,raw=T),data=s);summary(fit0Xsd)
fit0Zsd=lm(Zsd~poly(V,3,raw=T),data=s);summary(fit0Zsd)

## Store the sd and standardised data 
all$D65Xsd = predict(fit0Xsd,all)
all$D65Zsd = predict(fit0Zsd,all)
all$D65Xstd = all$D65Xoffset / all$D65Xsd
all$D65Zstd = all$D65Zoffset / all$D65Zsd

## Now model standardised responses 
fit4X = lm(D65Xstd~0+poly(C,3,raw=T)+C:cos((HA+16)/360*2*3.14159),data=all%>%filter(C<28,!(V>0&V<1))); summary(fit4X)
fit4Z = lm(D65Zstd~0+poly(C,2,raw=T)+C:sin((HA+13)/360*2*3.14159)+C:sin((HA-31)/180*2*3.14159),data=all%>%filter(C<28,!(V>0&V<1))); summary(fit4Z)

# Now model the INVERSE V operations from Y part of  X,Y,Z to V
fit9V = lm(V~poly(D65Y,7,raw=T),data=all%>%filter(C<28,!(V>0&V<1))); summary(fit9V)

# Now model the INVERSE C and HA operation from CIExy converted to polar co-ordinates
ggplot(data=all%>%filter(V>0.8),aes(x=CIEx-0.3127261,y=CIEy-0.3290336))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
all$a=27.35 + atan2(all$CIEy-0.3290336, all$CIEx-0.3127261)/(2*pi)*360    ##-180 to 180
all$m = sqrt((all$CIEx-0.3127261)*(all$CIEx-0.3127261)+(all$CIEy-0.3290336)*(all$CIEy-0.3290336))
all$dc = all$C / all$m     # ratio for r to C (ie C = r * dc)
all$da = all$HA - all$a    # difference from a to HA (ie HA = a + da)
all$da = all$da + (all$da > 180) * -360
all$VPred=predict(fit9V,all)
all$dv = all$V - all$VPred
ggplot(data=all%>%filter(C<32,V>0.8),aes(x=V,y=dv))+geom_point(aes(color=H))+geom_path(aes(group=V,alpha=0.1))

# First hemisphere Inverse for C needs adjusting depending on angle and V
ggplot(data=all%>%filter(a>0,C>0&C<32,V>.8),aes(x=a,y=dc))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9C1 = lm(dc~V*sin((a-4)/360*2*pi),data=all%>%filter(a>0,C>0,C<32)); summary(fit9C1)

# Second hemisphere, V>5 Inverse for C needs adjusting depending on angle and V
ggplot(data=all%>%filter(a<0,C>0&C<32),aes(x=a,y=dc))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9C2a = lm(dc~sin((a+77)/175*2*pi),data=all%>%filter(a<0,C>0&C<30,V>5&V<10)); summary(fit9C2a)

# Second hemisphere, V<=5 Inverse for C needs adjusting depending on angle and V
ggplot(data=all%>%filter(a<0,C>0&C<32,V<=5),aes(x=a,y=dc))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9C2b = lm(dc~poly(V,3,raw=T),data=all%>%filter(a<0,C>0&C<30,V<=5)); summary(fit9C2b)

# The residual angle appears to have relationship with a and V, (and C)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=a,y=da))+coord_polar()+geom_point(aes(color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~V)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~V)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=a,y=V))+geom_point(aes(color=H))+geom_path(aes(group=HA,alpha=0.1))+facet_wrap(~C)

# First quadrant then da has linear relationship with V, a and C
ggplot(data=all%>%filter(a>=90,C<32,V>0&V<10),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9HA1 = lm(da~V+a+C,data=all%>%filter(a>=90,C<32,V>0&V<10)); summary(fit9HA1)

# Second quadrant AND V>2 then da has linear relationship with V, a and C??
ggplot(data=all%>%filter(a>=0&a<90,C<32,V>0&V<10),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9HA2 = lm(da~V+a+C,data=all%>%filter(a>=0&a<90,C<32,V>2&V<10)); summary(fit9HA2)

# Third quadrant AND V>2 then da has linear relationship with a
ggplot(data=all%>%filter(a>= -90&a<0,C>0&C<32,V>0.8&V<10),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9HA3 = lm(da~a,data=all%>%filter(a>= -90&a<0,C>0&C<32,V>2&V<10)); summary(fit9HA3)

# Fouth quadrant then da has linear relationship with V, a and C
ggplot(data=all%>%filter(a< -90,C>0&C<32,V>0.8&V<10),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)
fit9HA4 = lm(da~V+a+C,data=all%>%filter(a< -90,C>0&C<32,V>0.8&V<10)); summary(fit9HA4)



# The prediction models for HA, V and C are approximate, use this as a seed for gradient descenet
# Store away the predicted inverse starting points
all$VPred=predict(fit9V,all)
all$CPred=predict(fit9C,all)
all$HAPred=27.35 + atan2(all$CIEy-0.3290336, all$CIEx-0.3127261)/(2*pi)*360 

# Need to standardise features for gradient descent
mean(all$HA); sd(all$HA); all$HAstd = (all$HA - mean(all$HA)) / sd(all$HA);
mean(all$V); sd(all$V); all$Vstd = (all$V - mean(all$V)) / sd(all$V);
mean(all$C); sd(all$C); all$Cstd = (all$C - mean(all$C)) / sd(all$C);


####################### test gradient descent

## Store away the standardised prediction and standard error
all$D65XstdPred=predict(fit4X,all)
all$D65ZstdPred=predict(fit4Z,all)
all$D65XstdSE=all$D65Xstd-all$D65XstdPred
all$D65ZstdSE=all$D65Zstd-all$D65ZstdPred

## Store away the prediction and standard error
all$D65XPred=all$D65Xmean+all$D65XstdPred*all$D65Xsd
all$D65ZPred=all$D65Zmean+all$D65ZstdPred*all$D65Zsd
all$D65YPred=all$D65Ymean

## Overwrite originals with predictions
all$D65X=all$D65XPred
all$D65Z=all$D65ZPred
all$D65Y=all$D65YPred
```

## Review the model
The following graphs compare predicted with actual data

```{r ReviewModel}
## Lets review the residual error
ggplot(data=all%>%filter(V>.8,C<30),aes(y=D65ZstdSE,x=HA,group=V,colour=V))+geom_point()+geom_line()+scale_x_continuous(breaks=seq(0, 360, by=360/8))+facet_wrap(~C,scales="free")

## lets review polar plots
ggplot(data=all%>%filter(V>-1),aes(y=D65X,x=HA,group=C,color=C))+coord_polar()+xlim(c(0,360))+geom_line()+geom_point(alpha=.6)+facet_wrap(~V)

## lets review D65X and D65Y together (showing paths with constant V,C,HA)
ggplot(data=all,aes(x=D65X,y=-D65Z))+geom_path(aes(group=V,color=V))+facet_wrap(~C)
ggplot(data=all%>%filter(V>0,C<30),aes(x=D65X,y=-D65Z))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~C,scales="free")
ggplot(data=all%>%filter(V==6),aes(x=D65X,y=-D65Z))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V,scales="free")

# how does D65X and D65Z vary relative to V,C and HA ***********************
ggplot(data=all%>%filter(V>.8,C>0),aes(x=C,y=D65Zoffset))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V,scales="free")
ggplot(data=all%>%filter(V>.8,C>0),aes(x=V,y=D65Zoffset))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~C,scales="free")


## ###########################################################
## ******* lets review CIEx and CIEy together (showing paths with constant V,C,HA)
ggplot(data=all%>%filter(V>0,C<30),aes(x=CIEx,y=CIEy))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~C,scales="free")
ggplot(data=all%>%filter(V>0.8),aes(x=CIEx,y=CIEy))+geom_point()+geom_path(aes(group=H,color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V)

## CIEx/CIEy delta angle = f(r,na,V)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=r,y=da))+geom_point(aes(color=V))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~na,scales="free")
## CIEx/CIEy delta angle = f(a, V, r)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=a,y=da))+geom_point(aes(color=H))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V,scales="free")
## C = f(H, r, and V)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=r,y=C))+geom_point(aes(color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~H,scales="free")
## CIEy vs CIEx with contour lines for equal hue (arcs) and equal chroma (ovals)
ggplot(data=all%>%filter(C>0,V>.8),aes(x=x,y=y))+geom_point(aes(color=H))+geom_path(aes(group=HA))+geom_path(aes(group=C,alpha=0.1))+facet_wrap(~V,scales="free")


```

## Creating the new dataset
Lets create a dataset for all Hues, all Values (0 to 10) and all Chromas (0 to 30)

```{r GenerateDataset}
## Generate a dataset for all Hues, all Values (0 to 10) and all Chromas (0 to 60)
alln <- all %>% filter(C==0) %>% select(h,V,C,H,HA,D65X,D65Y,D65Z)
for (h in levels(all$H)) {
  for (v in seq(1,9)) {
    new=data.frame(h=h,V=v,C=seq(2,24,by=2),H=h,HA=0,D65X=0,D65Y=0,D65Z=0)
    i=nrow(alln)
    alln[seq(i+1,i+nrow(new)),]=new[seq(1,nrow(new)),]
    }
  }

alln$HA <- (as.numeric(alln$H)-1)*360/40                    # create a continuous variable for modelling
alln$HA <- ifelse(is.na(alln$HA),-1,alln$HA)                  # set the HA to -1 for "N"

alln$D65XPred=predict(fit0X,alln)+predict(fit4X,alln)*predict(fit0Xsd,alln)
alln$D65ZPred=predict(fit0Z,alln)+predict(fit4Z,alln)*predict(fit0Zsd,alln)
alln$D65YPred=predict(fit0Y,alln)
alln$D65Xstd = predict(fit4X,alln)
alln$D65Zstd = predict(fit4Z,alln)
## Overwrite originals with predictions
alln$D65X=alln$D65XPred
alln$D65Z=alln$D65ZPred
alln$D65Y=alln$D65YPred

##all<-alln %>% filter(!(C>32),!((H>264)&(H<271)&(C>5+V*20.5/3.5)))

mun_XYZ_D65 <- matrix(c(all$D65X,all$D65Y,all$D65Z), ncol=3)


```

## Convert XYZ (D65) to Linear sRGB (D65)
Now we can convert from XYZ to Linear sRGB. The equation for conversion can be found at http://www.brucelindbloom.com/Eqn_RGB_XYZ_Matrix.html. We shall use the precalculated matrix for sRGB with D65 illumination.

```{r ConvertRGB1}
## sRGB profile transformation:
M_XYZ_to_sRGB_D65 <- matrix(c(3.24071, -0.969258, 0.0556352, -1.53726, 1.87599, -0.203996, -0.498571, 0.0415557, 1.05707), ncol=3, byrow=TRUE)

## apply the conversion matrix
mun_sRGB_D65 <- mun_XYZ_D65 %*% M_XYZ_to_sRGB_D65
```

## Convert Linear sRGB (D65) to sRGB (D65)
sRGB space is non-linear so we need to compand the linear RGB channels to a non-linear sRGB scale. The equations for this conversion can be found at http://www.brucelindbloom.com/Eqn_XYZ_to_RGB.html

```{r ConvertRGB2}

## define the transformation functions, these are applied on a conditional basis:
fun1 <- function(col_comp) { 1.055 * ( col_comp ^ ( 1 / 2.4 ) ) - 0.055 } 
fun2 <- function(col_comp) { 12.92 * col_comp } 

## the specific function is contingent on the absolute value of r,g,b components
all$R <- ifelse(mun_sRGB_D65[,1] > 0.0031308, fun1(mun_sRGB_D65[,1]), fun2(mun_sRGB_D65[,1]))
all$G <- ifelse(mun_sRGB_D65[,2] > 0.0031308, fun1(mun_sRGB_D65[,2]), fun2(mun_sRGB_D65[,2]))
all$B <- ifelse(mun_sRGB_D65[,3] > 0.0031308, fun1(mun_sRGB_D65[,3]), fun2(mun_sRGB_D65[,3]))

## create additional values based on these
all$hexRGB <- hex(sRGB(all$R,all$G,all$B))
all$LabL <- coords(as(sRGB(all$R,all$G,all$B), 'polarLAB'))[,1]
all$LabC <- coords(as(sRGB(all$R,all$G,all$B), 'polarLAB'))[,2]
all$LabH <- coords(as(sRGB(all$R,all$G,all$B), 'polarLAB'))[,3]

```

## Cleanup the data and clamp it to the sRGB space
Create a new data set that has all observations that are outside the sRGB space removed
Also clamp the RGB values from 0 to 1
```{r ClampRGB}
## create a data frame clipped to RGB scope
allc <- all %>% filter((R>= -0.001)&(R<=1.004),(G>= -0.004)&(G<=1.004),(B>= -0.004)&(B<=1.004),
                       (h=="N")|((V>=1)&(V<=9))) %>% arrange(HA)
allc$R <- pmax(0,pmin(allc$R,1))
allc$G <- pmax(0,pmin(allc$G,1))
allc$B <- pmax(0,pmin(allc$B,1))
allc$hexRGB <- hex(sRGB(allc$R,allc$G,allc$B))

ggplot(data=allc,aes(x=HA,y=C))+geom_point()+geom_path(aes(group=V))+facet_wrap(~V,scales="free")

## write the dataset out to a file
write.csv(data.frame(h=allc$h,V=allc$V,C=allc$C,hexRGB=allc$hexRGB), file = "/mnt/Media1/Projects/R-training/ColorSpace.csv")
```

## Plot all the tiles
Show all the munsell color tiles by hue, value and chroma

```{r DrawTiles, fig.width=15, fig.height=40}

ggplot(data = allc, aes(x = factor(C), y = factor(V))) +
    geom_tile(aes(fill = hexRGB), color="white", size = 1) +
    scale_colour_identity() +
    scale_x_discrete("Chroma") + 
    scale_y_discrete("Value", expand = c(0.125, 0)) +
    scale_fill_identity()+
    theme(aspect.ratio=1)+
    facet_wrap(~ H,ncol=4)

```
## Compare against other spaces
Show all the munsell colors against other color spaces

```{r DrawLab}
plot( as(sRGB(allc$R,allc$G,allc$B), 'HSV'), cex=0.5)
plot( as(sRGB(allc$R,allc$G,allc$B), 'LAB'), cex=0.5)
plot( as(sRGB(allc$R,allc$G,allc$B), 'polarLAB'), cex=0.5)
ggplot(aes(y=LabL,x=V,colour=V),data=all)+geom_line(aes(group=V))+geom_point()
ggplot(aes(y=LabH,x=H,colour=V),data=allc)+geom_point()
ggplot(aes(y=LabC,x=C,colour=V),data=allc)+geom_point()

## lets review x,y with the D65 midpoint shown
ggplot(data=all%>%filter(V>.8,CIEy>0,CIEy<2,CIEx<2,V>.8),aes(x=CIEx,y=CIEy))+geom_point(aes(shape=is.na(hexRGB)))+geom_vline(aes(xintercept=0.31271))+geom_hline(aes(yintercept=0.32902))+geom_path(aes(group=H,color=H))+geom_path(aes(group=C,alpha=0.1))+geom_polygon(data=all%>%filter(V>0.8,C>0,!is.na(hexRGB))%>%group_by(V,HA)%>%arrange(desc(C))%>%slice(1),alpha=.3)+facet_wrap(~V)
ggplot(data=all%>%filter(CIEy>0,CIEy<2,CIEx<2,V>.8),aes(x=CIEx,y=CIEy))+geom_point(aes(shape=is.na(hexRGB)))+geom_vline(aes(xintercept=0.31271))+geom_hline(aes(yintercept=0.32902))+geom_path(aes(group=H,color=H))+geom_path(aes(group=V,alpha=0.1))+facet_wrap(~C,scales="free") ## for a given level of C, the atan2(x,y) doesnt change much with V

```

## Show complement
Create a function to show the complement of a certain hue

```{r DrawComplement}
complement <- function(hue.name,  back.col = "white") {
  if (length(hue.name) > 1) stop("complement_slice currently only takes one hue")
  if (!hue.name %in% all$H) stop("invalid hue name")
  hues <- levels(all$H)
  index <- which(hues == hue.name)
  comp.hue <- hues[((index-1 + 20) %% 40)+1]
  all.sub <- subset(allc, H == hue.name | H == comp.hue)
  all.sub <- within(all.sub, { 
    C <- ifelse(H == comp.hue, -1, 1) * C
    H <- factor(H, levels = c(comp.hue, "N", hues[index])) 
    })
  
  ggplot(aes(x = C, y = V), data = all.sub) +
  geom_tile(aes(fill = hexRGB), colour = back.col,  size = 1) +
  scale_fill_identity() +
  scale_colour_identity() +
  scale_x_continuous("Chroma") +
  scale_y_continuous("Value") +
  facet_grid(. ~ H,  scales = "free_x", space = "free")  +
  coord_fixed()
  }
  
complement("2.5R")

```

(C) David Morrison 2015